---
title: "Reproducible Research: Peer Assessment 1"
output: html_document 
keep_md: true
---


``` {r LoadLibraries, echo = FALSE, include=FALSE}
library(dplyr)
library(tidyr)
```


## Loading and preprocessing the data

``` {r LoadData}

file <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"

if (!file.exists(file)){
        file <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
        download.file(file, "Tarea2.csv", method="curl")
}  

if (!file.exists("activity.csv")) { 
        unzip(file) 
}

activity <- read.csv("activity.csv")

head(activity)

```


## What is mean total number of steps taken per day?

Now we will calculate the total per day steps by summing the steps in each interval.
Next a histogram of this data is show

``` {r totalperday}

total.per.day <- activity %>% group_by(date) %>% summarise(total = sum(steps, na.rm = TRUE))

hist(total.per.day$total, main="Histogram of per day steps", xlab="Per day steps", col = "#80C610")

```

Having this information it will be calculated the mean and median of the total steps per day.
The mean is:

``` {r meanperday}
media <- mean(total.per.day$total)
print(media)

```


The median es:

``` {r medianperday}
mediana <- median(total.per.day$total)
print(mediana)
```


## What is the average daily activity pattern?

Next we will calculate the mean per interval of 5 minutes. 

``` {r meanperinterval}
mean.per.interval <- activity %>% group_by(interval) %>% summarise(average = mean(steps, na.rm = TRUE))

plot(mean.per.interval, type = "l", main="Average steps by interval", xlab="Interval", ylab="Step Average", col = "#80C687")
```

The maximum average of steps is in the interval:

``` {r maxmeanperinterval}

max.average <- which.max(mean.per.interval$average)

mean.per.interval$interval[max.average] 

```

## Imputing missing values

The number of NA's in the steps variable is:

``` {r countna}

n.na <- sum(is.na(activity$steps))

print(n.na)


```

For imputing values to NA we create a new data.frame call full.activity

``` {r newdataframe}
full.activity <- activity
```

Now we make a function that imputs the mean of the interval 

``` {r imputfunction}

for (i in 1:length(full.activity$steps)){
        if(is.na(full.activity$steps[i])){
                full.activity$steps[i]<-mean.per.interval$average[full.activity$interval[i]==mean.per.interval$interval]
        }
}

```

We check that there is no more missing values.

``` {r naimput}

sum(is.na(full.activity$steps))

```

With this new data frame we recalculate the total number of steps and make and histogram of this

``` {r totalperdayimput}

full.total.per.day <- full.activity %>% group_by(date) %>% summarise(total = sum(steps, na.rm = TRUE))

hist(full.total.per.day$total, col = "#80C687", main = "Histogram of total per day steps", xlab = "Per day steps")

```

Next we can see the mean of the imput data frame:

``` {r full mean}
full.media <- mean(full.total.per.day$total)
print(full.media)
```

The diference between the mean of the non imput with the imput data frame is:

``` {r full difmean}
dif.mean <- media - full.media
print(dif.mean)
```

The median of the imput data frame is:

``` {r full median}
full.mediana <- median(full.total.per.day$total)
print(full.mediana)
```

The diference between the median of the non imput with the imput data frame is:

``` {r full difmedian}
dif.median <- mediana - full.mediana
print(dif.median)
```

The imput generetes a disminution of the mean and median.

## Are there differences in activity patterns between weekdays and weekends?

Next it is created a new variable call clas.days that contains the option "weekday" or "weekend" depending the day of the week

``` {r }
full.activity$days<-weekdays(as.Date(full.activity$date))

full.activity$clas.days[full.activity$days=="lunes"]<-"Weekday"
full.activity$clas.days[full.activity$days=="martes"]<-"Weekday"
full.activity$clas.days[full.activity$days=="miércoles"]<-"Weekday"
full.activity$clas.days[full.activity$days=="jueves"]<-"Weekday"
full.activity$clas.days[full.activity$days=="viernes"]<-"Weekday"
full.activity$clas.days[full.activity$days=="sábado"]<-"Weekend"
full.activity$clas.days[full.activity$days=="domingo"]<-"Weekend"
```

It is calculated the mean per clas.days and interval

``` {r }
full.mean.per.interval <- full.activity %>% group_by(clas.days, interval) %>% summarise(average = mean(steps, na.rm = TRUE))
```

Last, we make a plot of the average steps by interval to see weekdays vs weekends

``` {r }
par(mfrow=c(1,1))

plot(full.mean.per.interval$average[full.mean.per.interval$clas.days=="Weekday"] ~ full.mean.per.interval$interval[full.mean.per.interval$clas.days=="Weekday"], type = "l", main="Average steps by interval by weekdays vs weekends", xlab="Interval", ylab="Step Average", col = "#FF0000")

lines(full.mean.per.interval$average[full.mean.per.interval$clas.days=="Weekend"]  ~ full.mean.per.interval$interval[full.mean.per.interval$clas.days=="Weekend"], type = "l", col = "#80FA10")

legend("topright",pch="--",  col = c("#FF0000", "#80FA10"), legend = c("Weekdays", "Weekends"))

```
